
<a id="Top"></a>
PCA - of the college stats data set as a whole and by decade
========================================================


*[SD/Decade](#SDdec)  
*[Loadings/Decade](#LOADdec)  
*[Each/Decade](#EACHdec)  


Principal Component Analysis as a whole

```{r}
qbStats = read.csv("F:/DataMining/project/data/qbStatsSubset.csv", header=TRUE)
test<- apply(qbStats,2,na.omit)
options(width=10000) 
# remove wins and touchdowns from this data set, b/c they are not predictors
qbStats = qbStats[,-4]
qbStats = qbStats[,-4]

qbStats = na.omit(qbStats)
b = scale(qbStats)
pcaP = princomp(b)
summary(pcaP)
abs(pcaP$loadings[,1:4])

```

Plots of PCA variances, scores, loadings and SD

```{r}
plot(pcaP)
plot(pcaP$scores, col = c("red", "blue"), pch = 19)
plot(pcaP$loadings, col = c("red", "blue"), pch = 19)
plot(pcaP$sdev)

```



```{r}
tens <- test[0:0]
zeros <- tens
nineties <- tens
eighties <- tens
seventies <- tens
sixties <- tens
fifties <- tens
forties <- tens
thirties <- tens

for(i in 1:nrow(qbStats)) {

  if(((qbStats$year[i])<=2014) && ((qbStats$year[i])>2009)) {
    tens <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=2009) && ((qbStats$year[i])>1999)) {
    zeros <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=1999) && ((qbStats$year[i])>1989)) {
    nineties <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=1989) && ((qbStats$year[i])>1979)) {
    eighties <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=1979) && ((qbStats$year[i])>1969)) {
    seventies <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=1969) && ((qbStats$year[i])>1959)) {
    sixties <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=1959) && ((qbStats$year[i])>1949)) {
    fifties <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=1949) && ((qbStats$year[i])>1939)) {
    forties <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }
  else if(((qbStats$year[i])<=1939) && ((qbStats$year[i])>1929)) {
    thirties <- rbind(as.data.frame(tens),as.data.frame(qbStats[i,]))
  }

}
# remove years, since we are comparing against decade
forties <- forties[,-5]
fifties <- fifties[,-5]
sixties <- sixties[,-5]
seventies <- seventies[,-5]
eighties <- eighties[,-5]
nineties <- nineties[,-5]
zeros <- zeros[,-5]
tens <- tens[,-5]

b = scale(tens)
pca10 = princomp(b)
summary(pca10)
abs(pca10$loadings[,1:4])

b = scale(zeros)
pca0 = princomp(b)
summary(pca0)
abs(pca0$loadings[,1:4])

b = scale(nineties)
pca9 = princomp(b)
summary(pca9)
abs(pca9$loadings[,1:4])

b = scale(eighties)
pca8 = princomp(b)
summary(pca8)
abs(pca8$loadings[,1:4])

b = scale(seventies)
pca7 = princomp(b)
summary(pca7)
abs(pca7$loadings[,1:4])

b = scale(sixties)
pca6 = princomp(b)
summary(pca6)
abs(pca6$loadings[,1:4])

b = scale(fifties)
pca5 = princomp(b)
summary(pca5)
abs(pca5$loadings[,1:4])

b = scale(forties)
pca4 = princomp(b)
summary(pca4)
abs(pca4$loadings[,1:4])

# c1a is component 1 for all the decades.
# the columns are the decades (1930's - 2010's)
# the rows are the features (age - c_avg_yds)
c1a <- matrix(nrow=10, ncol=8)
c2a <- matrix(nrow=10, ncol=8)
c3a <- matrix(nrow=10, ncol=8)
c4a <- matrix(nrow=10, ncol=8)

for(j in 1:10) {
  c1a[j,1] <- abs(pca4$loadings[j])
  c2a[j,2] <- abs(pca4$loadings[j+10])
  c3a[j,3] <- abs(pca4$loadings[j+20])
  c4a[j,4] <- abs(pca4$loadings[j+30])
}
for(j in 1:10) {
  c1a[j,2] <- abs(pca5$loadings[j])
  c2a[j,3] <- abs(pca5$loadings[j+10])
  c3a[j,4] <- abs(pca5$loadings[j+20])
  c4a[j,5] <- abs(pca5$loadings[j+30])
}
for(j in 1:10) {
  c1a[j,3] <- abs(pca6$loadings[j])
  c2a[j,4] <- abs(pca6$loadings[j+10])
  c3a[j,5] <- abs(pca6$loadings[j+20])
  c4a[j,6] <- abs(pca6$loadings[j+30])
}
for(j in 1:10) {
  c1a[j,4] <- abs(pca7$loadings[j])
  c2a[j,5] <- abs(pca7$loadings[j+10])
  c3a[j,6] <- abs(pca7$loadings[j+20])
  c4a[j,7] <- abs(pca7$loadings[j+30])
}
for(j in 1:10) {
  c1a[j,5] <- abs(pca8$loadings[j])
  c2a[j,6] <- abs(pca8$loadings[j+10])
  c3a[j,7] <- abs(pca8$loadings[j+20])
  c4a[j,8] <- abs(pca8$loadings[j+30])
}
for(j in 1:10) {
  c1a[j,6] <- abs(pca9$loadings[j])
  c2a[j,7] <- abs(pca9$loadings[j+10])
  c3a[j,8] <- abs(pca9$loadings[j+20])
  c4a[j,1] <- abs(pca9$loadings[j+30])
}
for(j in 1:10) {
  c1a[j,7] <- abs(pca0$loadings[j])
  c2a[j,8] <- abs(pca0$loadings[j+10])
  c3a[j,1] <- abs(pca0$loadings[j+20])
  c4a[j,2] <- abs(pca0$loadings[j+30])
}
for(j in 1:10) {
  c1a[j,8] <- abs(pca10$loadings[j])
  c2a[j,1] <- abs(pca10$loadings[j+10])
  c3a[j,2] <- abs(pca10$loadings[j+20])
  c4a[j,3] <- abs(pca10$loadings[j+30])
}

factorNames <- colnames(sixties)

# std deviations of components
sda <- matrix(nrow=10, ncol=8)

for(j in 1:10) {
  sda[j,1] <- pca4$sdev[j]
  sda[j,2] <- pca5$sdev[j]
  sda[j,3] <- pca6$sdev[j]
  sda[j,4] <- pca7$sdev[j]
  sda[j,5] <- pca8$sdev[j]
  sda[j,6] <- pca9$sdev[j]
  sda[j,7] <- pca0$sdev[j]
  sda[j,8] <- pca10$sdev[j]
}
print(sda)
print(summary(pca4))
print(summary(pca5))
print(summary(pca6))
print(summary(pca10))
```

```{r}
```

<a id="SDdec"></a>
Plots of the Standard Deviations of the first 4 PCs over the decades...
========================================================
[Top](#Top)  


```{r}
```

```{r}
components <- c("PC1", "PC2", "PC3", "PC4")
for(h in 1:4){
  plot(sda[h,], type="o", main=components[h], sub="Standard Deviations over decades", xlab="decades (1=1930's, 8=2010's)",ylab="Std Deviation", col="blue")
}


```

```{r}
```


<a id="LOADdec"></a>
Plots of the loadings of the first 4 PCs over the decades...
========================================================
[Top](#Top)  



```{r}
```

```{r}
for(h in 1:10){
  plot(c1a[h,], type="o", main=" PC1 ", sub="loading over decades", xlab="decades (1=1930's, 8=2010's)",ylab=factorNames[h], col="blue")
}
for(h in 1:10){
  plot(c2a[h,], type="o", main=" PC2 ", sub="loading over decades", xlab="decades (1=1930's, 8=2010's)",ylab=factorNames[h], col="blue")
}
for(h in 1:10){
  plot(c3a[h,], type="o", main=" PC3 ", sub="loading over decades", xlab="decades (1=1930's, 8=2010's)",ylab=factorNames[h], col="blue")
}
for(h in 1:10){
  plot(c4a[h,], type="o", main=" PC4 ", sub="loading over decades", xlab="decades (1=1930's, 8=2010's)",ylab=factorNames[h], col="blue")
}

```

```{r}
```


<a id="EACHdec"></a>
Plots of PCA, scores and loading for each decade...
========================================================
[Top](#Top)  


```{r}
```

```{r fig.width=10, fig.height=10}
plot(pca10)
plot(pca10$scores, col = c("red", "blue"), pch = 19)
plot(pca10$loadings, col = c("red", "blue"), pch = 19)

plot(pca0)
plot(pca0$scores, col = c("red", "blue"), pch = 19)
plot(pca0$loadings, col = c("red", "blue"), pch = 19)

plot(pca9)
plot(pca9$scores, col = c("red", "blue"), pch = 19)
plot(pca9$loadings, col = c("red", "blue"), pch = 19)

plot(pca8)
plot(pca8$scores, col = c("red", "blue"), pch = 19)
plot(pca8$loadings, col = c("red", "blue"), pch = 19)

plot(pca7)
plot(pca7$scores, col = c("red", "blue"), pch = 19)
plot(pca7$loadings, col = c("red", "blue"), pch = 19)

plot(pca6)
plot(pca6$scores, col = c("red", "blue"), pch = 19)
plot(pca6$loadings, col = c("red", "blue"), pch = 19)

plot(pca5)
plot(pca5$scores, col = c("red", "blue"), pch = 19)
plot(pca5$loadings, col = c("red", "blue"), pch = 19)

plot(pca4)
plot(pca4$scores, col = c("red", "blue"), pch = 19)
plot(pca4$loadings, col = c("red", "blue"), pch = 19)



```


[Top](#Top)  